---
title: "RAPID Fire Project - Lake Profile Data"
date: "2022-10-03"
author: "J. Brentrup"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Lakes are cool
good place for notes
this section of code will load my library

# Load Libraries
```{r, message=FALSE}

# install.packages("tidyverse")

library(tidyverse)
library(googledrive)
library(readxl)
library(lubridate)
```

# Download profile data
- load data directly from google drive folder
```{r, message=FALSE}

# See what files are available
drive_find(q = "name contains 'Profiles'")

# Download Lake Profiles Round 1

# Use file ID
# drive_download(as_id("13qx6D_9t2AW55z2Oqkfna_zNLHowxvh_"),
#                path = "Data/Field Sampling_Lake Profiles_Round1.xlsx", 
#                overwrite = TRUE)

drive_download("Field Sampling_Lake Profiles_Round1.xlsx",
               path = "Data/WaterQuality/Field Sampling_Lake Profiles_Round1.xlsx", 
               overwrite = TRUE)

# Download Lake Profiles Round 2
drive_download("Field Sampling_Lake Profiles_Round2.xlsx",
               path = "Data/WaterQuality/Field Sampling_Lake Profiles_Round2.xlsx", 
               overwrite = TRUE)

```

# Load profile data - full time complete
```{r, warning=FALSE}

profiles_all <- read_csv("Data/WaterQuality/Field-Sampling_Lake-Profiles10-19-22Complete.csv", col_types = c(
  Site = col_character(),
  Date = col_character(),
  ProfileDepth_m = col_double(),
  WaterTemp_C = col_double(),
  SpecCond_uScm = col_double(),
  LDO_mgL = col_double(),
  LDO_pct = col_double(),
  pH = col_double()
  
))

str(profiles_all)
# Convert date to posixt for

profiles_all$Date <- as_date(profiles_all$Date, format = " %m/%d/%Y")

# add month
profiles_all <- profiles_all %>% 
  mutate(Month = month(Date))

```

# Check profile data
```{r}
summary(profiles_all)

# Fix high DO in Eighteen lake
which.max(profiles_all$LDO_mgL) # high value for eighteen lake DO concentration and %Sat switched in July

# Fix high DO values- got switched % Sat and for mg/L
profiles_all[361,6] <- 9.02
profiles_all[361,7] <- 108.9

profiles_all[362,6] <- 9.18
profiles_all[362,7] <- 108.8

profiles_all[363,6] <- 9.2
profiles_all[363,7] <- 109.7

profiles_all[364,6] <- 9.34
profiles_all[364,7] <- 111.5



# Total number of lakes - 30
unique(profiles_all$Site)

#Filter for lakes > 2m - 14 lakes

deep_lakes <- profiles_all %>% 
  filter(ProfileDepth_m > 2) %>% 
  count(Site)

# Filter for deep lakes (profiles > 2 m)
deep_lakes_profiles <- profiles_all %>% 
   filter(Site %in% deep_lakes$Site)

```

# Plot water temp profiles for deep lakes
```{r}

ggplot(data = deep_lakes_profiles, aes(x = WaterTemp_C, y = ProfileDepth_m, color = factor(Month))) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()+
  facet_wrap(~Site)

```


# Download and load latest water quality data
```{r}
# drive_download("Lab Analyses_Lakes9-30-22.xlsx",
#                path = "Data/WaterQuality/LabAnalyses_Lakes9-30-22.xlsx", 
#                overwrite = TRUE)

wq_data <- read_csv("Data/WaterQuality/combined_lab_field_may_sep.csv")
# str(wq_data)

str(wq_data)
summary(wq_data)

which.max(wq_data$LDO_mgL) # high value for eighteen lake DO concentration and %Sat switched in July

# Fix high DO values- got switched 108.9 for % Sat and 9.02 for mg/L
wq_data[23,36] <- 9.02
wq_data[23,37] <- 108.9

# Rename depth column in WQ to separate from profile depth
wq_data <- wq_data %>% 
  rename(Depth_m = ProfileDepth_m)


```


# Join profile and WQ data 
- match by site name and month to get just lake characteristic attributes (burn, connectivity class)
```{r}

profiles_wq_data <- full_join(profiles_all, wq_data[,c(1:2, 16, 18:19)], by = c("Site", "Month"))

summary(profiles_wq_data)

```


# Plot profiles
```{r}

# Deepest lake - 4th McDougal only in May
deep_lake <- deep_lakes_profiles_all %>% 
  filter(Site == "Fourth McDougal")

ggplot(data = deep_lake, aes(x = WaterTemp_C, y = ProfileDepth_m)) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()

# Filter for deep lakes (profiles > 2 m)
deep_lakes_profiles_all <- profiles_wq_data %>% 
   filter(Site %in% deep_lakes$Site) %>% 
  filter(!is.na(Month))

# Water Temp
ggplot(data = deep_lakes_profiles_all, aes(x = WaterTemp_C, y = ProfileDepth_m, color = Group, shape = factor(Month))) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()+
  labs(color = "Lake Type", shape = "Month")+
  facet_wrap(~Site)

# DO mg/L
names(deep_lakes_profiles_all)

ggplot(data = deep_lakes_profiles_all, aes(x = LDO_mgL, y = ProfileDepth_m, color = Group, shape = factor(Month))) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()+
  labs(color = "Lake Type", shape = "Month")+
  facet_wrap(~Site)

# DO % Sat
ggplot(data = deep_lakes_profiles_all, aes(x = LDO_pct, y = ProfileDepth_m, color = Group, shape = factor(Month))) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()+
  labs(color = "Lake Type", shape = "Month")+
  facet_wrap(~Site)

# pH
ggplot(data = deep_lakes_profiles_all, aes(x = pH, y = ProfileDepth_m, color = Group, shape = factor(Month))) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()+
  labs(color = "Lake Type", shape = "Month")+
  facet_wrap(~Site)

# specific conductivity
ggplot(data = deep_lakes_profiles_all, aes(x = SpecCond_uScm, y = ProfileDepth_m, color = Group, shape = factor(Month))) +
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  geom_point()+
  geom_path()+
  labs(color = "Lake Type", shape = "Month")+
  facet_wrap(~Site)

```

# Summary
```{r}
burn_lakes <- deep_lakes_profiles_all %>% 
  group_by(Group, Month) %>% 
  summarize(mean = mean(LDO_mgL, na.rm = T), min = min(LDO_mgL, na.rm = T), max = max(LDO_mgL, na.rm = T), sd = sd(LDO_mgL, na.rm = T)) 

```

